---
title: "Make 3D maps"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: myStyle.css
    highlight: haddock
    toc: true
    toc_float: true
    number_sections: false
    code_folding: show
    df_print: paged
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  message = FALSE, # prevents messages that are generated by code from appearing in the finished file
  warning = FALSE, # prevents warnings that are generated by code from appearing in the finished
  error = TRUE, # if FALSE, knitr will not display any error messages generated by the code
  eval = TRUE, # if FALSE, knitr will not run the code in the code chunk
  include = TRUE, # f FALSE, knitr will run the chunk but not include the chunk in the final document
  results = 'hide', # If 'hide', knitr will not display the code's results in the final document. If 'hold', knitr will delay displaying all output pieces until the end of the chunk. If 'asis', knitr will pass through results without reformatting them (useful if results return raw HTML, etc.)
  fig.show = 'asis', # If 'hide', knitr will generate the plots created in the chunk, but not include them in the final document. If 'hold', knitr will delay displaying the plots created by the chunk until the end of the chunk. 
  fig.keep = 'last',
  fig.align = 'center',
  fig.width = 4.5,
  fig.height = 3,
  max.print = 3, # max no. of lines in data frames
  strip.white = TRUE, # remove leading and lagging white lines
  comment = NA # to remove all hashes
  )  
```

# Preliminaries

```{r}
# devtools::install_github('tylermorganwall/rayshader')
pacman::p_load(rayshader)
pacman::p_load(rgdal)
pacman::p_load(proj4)
pacman::p_load(raster)
pacman::p_load(geoviz)
pacman::p_load(shiny)
pacman::p_load(leaflet)
pacman::p_load(geosphere)
pacman::p_load(scales)

sapply(list.files(here::here('R')), function(x) source(here::here('R', x)))

# Gijón
boundingBox <- DefineBoundingBox(p1 = c(lon = -5.75, lat = 43.45), p2 = c(lon = -5.55, lat = 43.6))
```


# Obtain georeferenced tif

https://www.qgistutorials.com/en/docs/3/georeferencing_basics.html

Using QGIS 2.18:

- Open the map used as reference to get the coordinates to georeference a .jpg image: `Layer > Add Layer > Add Raster Layer` (Ctr+Shift+R). Select the file `Asturias_raster.ecw`.

- Select projection `+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs` in `Project > Project Properties > Enable 'On The Fly' SRC = WGS 84 / UTM zone 30N EPSG:32360`.

- Activate plugin: `Plugins > Manage and Install plugins... > Install > Georeferencer GDAL`.

- Open Georeferencer: `Raster > Georeferencer > Georeferencer`.

- Open .jpg: `File > Open Raster...` (Ctrl+O).

- Add at least six points: `Edit > Add Point`. Use the option `From canvas map` to select the same geographic point in the reference raster map (`Asturias_raster.ecw`).

- Define the transformation settings: `Settings > Transformation Settings`. `Transformation = Polynomial 1`, `Rsampling method = Nearest neighbour`, `Target CRS = EPSG:32360 - ETRS89/ETRS-TM30`, `Compression = LZW`. 

# Manually define the bounding box

http://archived.mhermans.net/hiking-gpx-r-leaflet.html


```{r}
rm(boundingBox)

SelectMapArea()

ShowOSM(boundingBox, graticuleInterval = 0.1)
```


_____________________________________

# TIF data

Load geoTIF:

```{r}
folderName <- '\\\\pocpaco\\maps\\overlays\\'
tifData_raster <- raster::stack(paste(folderName, 'Asturias_espacios_protegidos_georef.tif', sep = '\\'))

# How many bands?
cat(sprintf('Number layers found in the TIF file: %d:\n ', nlayers(tifData_raster)))

# Which coordinate reference system?
cat(sprintf('TIF CRS: %s\n', raster::crs(tifData_raster)))
```

Crop:

```{r}
tifData_raster <- tifData_raster %>% 
  raster::crop(c(boundingBox$p1$x, boundingBox$p2$x, boundingBox$p1$y, boundingBox$p2$y))
```


Plot:

```{r}
raster::plotRGB(tifData_raster)
```


Get borders using shape file:

```{r}
mapShape <- rgdal::readOGR('\\\\pocpaco\\maps\\shapes\\recintos_autonomicas_inspire_peninbal_etrs89.shp') %>% 
  subset(NAMEUNIT == 'Principado de Asturias') %>% 
  sp::spTransform(CRSobj = crs(tifData_raster)) %T>% 
  plot()
```


Crop:

Mask: https://stackoverflow.com/questions/23073669/clipping-raster-using-shapefile-in-r-but-keeping-the-geometry-of-the-shapefile

```{r}
croppedRaster <- tifData_raster %>% 
  raster::crop(extent(mapShape)) %>% 
  raster::mask(mapShape)

```




Transform:

```{r}
## Transform each band to array
tifData_matrix_1 <- rayshader::raster_to_matrix(tifData_raster[[1]])
tifData_matrix_2 <- rayshader::raster_to_matrix(tifData_raster[[2]])
tifData_matrix_3 <- rayshader::raster_to_matrix(tifData_raster[[3]])

## Merge band arrays into one tensor:
### Initialize
tifData_matrix_all <- array(0, dim = c(nrow(tifData_matrix_1), ncol(tifData_matrix_1), 3))

### Merge
tifData_matrix_all[, , 1] <- tifData_matrix_1/255 
tifData_matrix_all[, , 2] <- tifData_matrix_2/255 
tifData_matrix_all[, , 3] <- tifData_matrix_3/255

### Transpose to abide by the elevation raster orientation
tifData_matrix_all <- aperm(tifData_matrix_all, c(2, 1, 3))

### Plot
plot_map(tifData_matrix_all)
```



# Elevation data

Load:

```{r}
elevationData_raster <- readRDS(here::here('output', 'rasterObject_Asturias.rds'))

# Which coordinate reference system?
raster::crs(elevationData_raster)

```

```{r}
elevationData_raster <- elevationData_raster %>% 
  raster::crop(extent(mapShape)) %>% 
  raster::mask(mapShape)

```

Transform the elevation data from to the same coordinate system as the tiff:

```{r}
elevationData_raster <- elevationData_raster %>% 
  raster::crop(c(boundingBox$p1$x, boundingBox$p2$x, boundingBox$p1$y, boundingBox$p2$y)) %T>% 
  plot()
```

```{r}
elevationData_matrix <- rayshader::raster_to_matrix(elevationData_raster)

# Remove NAs (ocean)
elevationData_matrix[is.na(elevationData_matrix)] <- 0
```


# Plot

```{r}
plot_3d(tifData_matrix_all, elevationData_matrix, windowsize = c(1100,900), zscale = 50, shadowdepth = -50,
        zoom=0.5, phi=45,theta=-45,fov=70, background = "#F2E1D0", shadowcolor = "#523E2B")

render_snapshot(title_text = '',
                title_bar_color = "#1f5214", title_color = "white", title_bar_alpha = 1)
```






